# Process, clean, and reshape data

```{r}
#| label: packages-data
#| warning: false
#| message: false

library(tidyverse)
```

Popular conjoint survey platforms like Qualtrics and Sawtooth typically provide results data in two separate data files: (1) individual participant-level responses and (2) a bank of possible alternative-level combinations of features. To analyze the results of a conjoint experiment, the two datasets need to be joined.

## Individual-level responses

```{r}
responses <- readRDS(here::here("data", "processed_data", "responses_illustration.rds"))
responses
```


## Possible alternatives

```{r}
alternatives <- readRDS(here::here("data", "processed_data", "alternatives_illustration.rds"))
alternatives
```

```{r}
alternatives |> 
  summarize(
    versions = n_distinct(version),
    questions = n_distinct(question),
    alts = n_distinct(alt)
  )
```

## Pivoting and expanding

The original responses data is wide, with a column for each of the 12 choices. We first need to make it long, with a row for each respondent-choice

```{r}
responses_long <- responses |> 
  pivot_longer(
    cols = starts_with("CBC_Random"),
    names_to = "question_raw",
    values_to = "chosen_alt"
  ) %>%
  # The task number is embedded in text, like "CBC_Random6"; this extracts it
  mutate(question = as.numeric(str_extract(question_raw, "\\d+"))) %>%
  select(-question_raw)
responses_long
```

```{r}
responses_long |> 
  summarize(
    resp_ids = n_distinct(resp_id),
    questions = n_distinct(question)
  )
```

```{r}
responses_long_expanded <- responses_long |>
  expand(resp_id, question, alt = 1:2) |> 
  left_join(responses_long, by = join_by(resp_id, question))
responses_long_expanded
```

```{r}
responses_long_expanded |> 
  summarize(
    resp_ids = n_distinct(resp_id),
    questions = n_distinct(question),
    alts = n_distinct(alt)
  )
```

## Final data

```{r}
combined <- responses_long_expanded |> 
  left_join(alternatives, by = join_by(resp_id == version, question, alt)) |> 
  mutate(choice = as.numeric(alt == chosen_alt))
```

```{r}
nrow(combined) / 2 / 12
```
