---
title: "The Conjoint Cookbook"
format: 
  gfm: 
    wrap: preserve
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

library(emoji)
```

[Andrew Heiss](https://www.andrewheiss.com/) • Andrew Young School of Policy Studies • Georgia State University

## `r emoji_glue(":1st_place_medal::gorilla:")`: Note on "golden gorilla" project name

Because project titles change all the time with revisions, rewriting, and peer review, I used [{codename}](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames) internal-to-me project name that won’t change.

```{r generate-code-name, eval=FALSE}
library(codename)
codename_message()
#> code name generated by {codename} v.0.5.0. R version 4.4.1 (2024-06-14).

codename(seed = "conjoint psych", type = "ubuntu")
#> [1] "golden gorilla"
```

## Colors

The main red color (`#F2274C`) was taken from a standard red/white apron image I found on Google Docs. The other colors are roughly perceptually balanced using HCL hues around that red:

```
#2E5BE4  (blue)
#1AB07A  (green-teal)
#F2274C  (red)
#F9A825  (gold)
#8E44AD  (purple)
```

## Icons in logo

The icons in the fake conjoint survey in the logo come from [Streamline](https://home.streamlinehq.com/)'s [Core Solid set](https://www.streamlinehq.com/icons/core-solid-free), and are Creative Commons licensed:

- [Fork Spoon](https://www.streamlinehq.com/icons/download/fork-spoon--26756)
- [Fork Knife](https://www.streamlinehq.com/icons/download/fork-knife--26756)
- [Burger](https://www.streamlinehq.com/icons/download/burger--26756)
- [Toast](https://www.streamlinehq.com/icons/download/toast--26756)
- [Ice Cream 3](https://www.streamlinehq.com/icons/download/ice-cream-3--26756)
- [Ice Cream 2](https://www.streamlinehq.com/icons/download/ice-cream-2--26756)
